---
title: "effects"
author: "Rich"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file visualises and tabulates parameters from models generated
from the models.Rmd script.

# load the libraries that we will be using #

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "brms", 
         "tidybayes", "bayesplot", "future", "parallel")

lapply(pkg, library, character.only = TRUE)
```

# plot settings #

theme settings for ggplot

```{r ggplot-settings, inlcude = F}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 18), 
          title = element_text(size = 18),
          legend.position = "bottom")
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# read in previously saved brms models object files and/or fits ##

If available, this saves time computing them each time 

```{r read-models}
# b1 <- readRDS("exp1/models/b1.rds") # dv is centred and standardised
b2 <- readRDS("exp1/models/b2.rds") # dv is in original units
```

# Look at the point and interval estimates for model b1 #

using posterior_samples and tidybayes

```{r posterior-samples}
post <- as_draws_df(b2) %>%  #
  select(-lp__) %>% 
  as_tibble()

post_summary <- posterior_summary(b2, robust = TRUE)

# all fixed effects
post_qi_b <- post %>%
  select(starts_with("b_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(post_qi_b)

# varying effects - SD
post_qi_sd <- post %>%
  select(starts_with("sd_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(post_qi_sd)

# individual item and SD variation
# all individual pt variation
vary_vals <- post %>%
  select(starts_with("r_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(vary_vals)
```

## Visualise fixed effects ##

Using posterior_samples and tidybayes

```{r fixed-effects-plot}
## fixed effects of interest, excluding basic intercepts for the ordinal model 
tidy_fixed <- post %>% 
  select(starts_with("b_")) %>% # select and rename in simpler labels
  pivot_longer(everything()) %>% # move from wide to long
  mutate(key = factor(name, levels=unique(name)),
         term = if_else(str_detect(key, "stimd:compatd"), "interaction",
                if_else(str_detect(key, "compatd"), "compatibility",
                if_else(str_detect(key, "stimd"), "stimulus", "intercept"))),
         term = factor(term, levels=c("intercept", "stimulus", "compatibility",
                                      "interaction")))  
head(tidy_fixed)
tail(tidy_fixed)

check.labels <- tidy_fixed %>% 
 group_by(key, term) %>% 
 distinct(key, term)
check.labels

## for model b2, remove intercept to zoom in on effects if interest
tidy_fixed <- tidy_fixed %>% 
  filter(term != "intercept")
head(tidy_fixed)
str(tidy_fixed)

p_tidy_fixed <- ggplot(tidy_fixed, aes(x = value, y = fct_rev(term), fill=term)) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye(point_size = 1) +
  labs(title = "Coefficient plot for fixed effects (predictors)",
       x = NULL, y = NULL) +
  scale_fill_brewer(palette="Dark2") +
  theme(panel.grid   = element_blank(),
        panel.grid.major.y = element_line(color = alpha("firebrick4", 1/2), linetype = 3),
        axis.text.y  = element_text(hjust = 0),
        axis.ticks.y = element_blank(),
        legend.position = "none") +
  labs(x= "reaction time (ms)") +
  coord_cartesian(xlim =c(-20, 25))+
  scale_x_continuous(breaks=seq(-20, 25, 5))
p_tidy_fixed

ggsave ("exp1/figures/b2_fixef.png",
        width = 6, height = 8)
```

## make a table of fixed effects ##

```{r fixed-effects-table}
fix_tab <- post_qi_b %>%
  mutate(name = factor(name, levels=unique(name)),
         term = if_else(str_detect(name, "stimd:compatd"), "interaction",
                if_else(str_detect(name, "compatd"), "compatibility",
                if_else(str_detect(name, "stimd"), "stimulus", "intercept"))),
         term = factor(term, levels=c("intercept", "stimulus", "compatibility",
                                      "interaction"))) %>% 
  arrange(term) %>% 
  select(term, value, .lower, .upper)
fix_tab

write_csv(fix_tab, "exp1/tables/b2_fix_tab.csv")
# this table data can be read in when generating the manuscript and papaja can produce the table formatting etc.
```

## Visualise varying effects ##

using posterior samples and tidybayes.

## Variance at the group level (i.e., intercepts and effects across participants) ##

```{r varying-effects}
## varying effects of interest, including item, as well as participant and effects across pts
tidy_vary <- post %>% 
  select(contains("sd_")) %>% 
  pivot_longer(everything()) %>% 
  mutate(key = factor(name, levels=unique(name)),
         term = if_else(str_detect(key, "stimd:compatd"), "interaction",
                if_else(str_detect(key, "compatd"), "compatibility",
                if_else(str_detect(key, "stimd"), "stimulus", "intercept"))),
         term = factor(term, levels=c("intercept", "stimulus", "compatibility",
                                      "interaction"))) %>% 
  arrange(term)
head(tidy_vary)
tail(tidy_vary)

check.vary <- tidy_vary %>% 
  group_by(key, term) %>% 
  distinct(key, term)
check.vary

p_tidy_vary <- ggplot(tidy_vary, aes(x = value, y = fct_rev(term), fill=term)) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye() +
  labs(title = "SD plot for varying effects",
       x = NULL, y = NULL) +
  scale_fill_brewer(palette="Dark2") +
  theme(panel.grid   = element_blank(),
        panel.grid.major.y = element_line(color = alpha("firebrick4", 1/2), linetype = 3),
        # axis.text.y  = element_blank(),
        # axis.ticks.y = element_blank(),
        legend.position = "none") 
p_tidy_vary

ggsave ("exp1/figures/b2_vary.png",
        width = 6, height = 8)
```

## Variance within group levels i.e., at the individual item and participant level ##

Note - this should be in a centred metric, so that deviation from the group average (zero) can be visualised.

## filter and wrangle varying effects within participants ##

```{r vary-pid}
# 102 pts x 4 terms = 408 params
tidy_varyvals <- post_summary %>%
  data.frame() %>% 
  rownames_to_column("param") %>% 
  filter(str_detect(param, "r_pid"),
         str_detect(param, "cor", negate = TRUE)) %>% 
  mutate(term = if_else(str_detect(param, "stimd:compatd"), "interaction",
                if_else(str_detect(param, "compatd"), "compatibility",
                if_else(str_detect(param, "stimd"), "stimulus", "intercept"))),
         term = factor(term, levels=c("intercept", "stimulus", "compatibility",
                                      "interaction"))) %>% 
  arrange(term, Estimate) %>% 
  mutate(rank = rep(1:102, times=4),
         outlier = if_else(Q2.5 > 0 | Q97.5 < 0, "yes", "no"), 
         outlier = factor(outlier, levels = c("no", "yes")))
head(tidy_varyvals)

check.vary.pt <- tidy_varyvals %>%
  group_by(term) %>% 
  tally()
check.vary.pt
```

## plot varying effects via caterpillar plots ##

```{r plot-vary-pid}
p_tidy_varyvals <- ggplot(tidy_varyvals, aes(x=rank, y=Estimate, colour=outlier))+
  geom_pointrange(aes(ymin=Q2.5,ymax=Q97.5),fatten = 0.5)+
  geom_hline(yintercept=0, colour = "red")+
  scale_colour_manual(values=c("darkgrey","black")) +
  labs(title = "Varying effects by participant",
       x = "participant")+
  theme(axis.text.x  = element_blank())+
  guides(colour = "none")+
  facet_grid(~term)
p_tidy_varyvals

ggsave ("exp1/figures/b2_varyvals.png",
        width = 6, height = 8)
```



