---
title: "wrangle"
author: "Rich"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file wrangles raw data from an old gaze-cueing experiment that Andrew and
some masters students completed.

# load the libraries that we will be using #

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "ggridges")

lapply(pkg, library, character.only = TRUE)
```

# plot settings #

theme settings for ggplot

```{r ggplot-settings, inlcude = F}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 18), 
          title = element_text(size = 18),
          legend.position = "bottom")
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# section 1 #

## read in the raw data ##

```{r read-data}
raw <- read_csv("exp1/data/data.csv")
head(raw)
tail(raw)
glimpse(raw)
```

# section 2 #

## do some wrangling ##

recode pid

```{r recode-pid}
data <- raw %>%
  mutate(pid = (str_remove(pid, "[pP]")),
         pid = as.numeric(pid),
         pid = if_else(is.na(pid), 2, pid)) # this is a bit hacky
head(data)

check.pid <- data %>% 
  distinct(pid)
check.pid
```

recode stimuli

```{r recode-stimuli}
# check
check.stim <- data %>%
  distinct(stim)
check.stim

## simplify stim coding by turning all gaze ids to be the same
data <- data %>%
  mutate(stim = if_else(str_detect(stim, "stim"), "gaze", stim)) 
head(data)
str(data)
```

remove unnecessary columns and create factors

```{r create-factors}
data <- data %>%
  select(-gaze, -target, -isi) %>% 
  mutate(stim = factor(stim,
                       levels = c("arrow", "gaze")),
         compat = factor(compat, 
                         levels = c("compat", "incomp")),
         rt = as.numeric(rt*1000),
         acc = as.numeric(acc))
head(data)
str(data)
```


## section 3 ##

## make some summary data ##

## accuracy ##

summary per pid

```{r acc-pid}
acc_pid <- data %>% 
  group_by(pid, stim, compat) %>% 
  summarise(n = n(), 
            sum = sum(acc), 
            perc = sum/n*100)
acc_pid
```

group summary

```{r acc-group}
acc_group <- acc_pid %>% 
  group_by(stim, compat) %>% 
  summarise(n = length(unique(pid)),
            mean_sum = mean(sum), 
            mean_perc = mean(perc),
            sd = sd(perc),
            sem = sd/sqrt(n)) %>% 
  rename(sum = mean_sum, perc = mean_perc)
acc_group
```

plot accuracy 

```{r plot3.1}
p3.1 <- ggplot(acc_pid, 
               aes(x = compat, y = perc, fill = compat)) +
  geom_jitter(position=position_jitterdodge(dodge.width =1), 
               alpha = 1, colour = "darkgrey") +
  geom_violin(alpha = 0.7, position=pd2) +
  geom_point(data = acc_group, size = 3) +
  geom_errorbar(data = acc_group,
                aes(y = perc, ymin = perc-sem, ymax = perc+sem),
                width=.2, position=pd2) +
  scale_fill_brewer(palette = "Dark2") +
  geom_line(data = acc_group, aes(y = perc, group=1)) +
  facet_wrap(~stim) +
  ylim(80, 100) +
  labs(title = "Accuracy")
p3.1   

ggsave("exp1/figures/acc_violin.png",
       width = 6, height = 8)
```

acc is very high - 95-96% for all conditions.
slightly more accurate in the compat than incomp, but the difference is tiny <0.5% diff. 
That means we can strip errors from the dataset and focus on RT. 

## rt ##

first, filter errors (i.e., where they pressed the wrong button).

```{r filter-acc}
fdata <- data %>% 
  filter(acc == 1)
head(fdata)
```

create summary data

summary per pid

```{r rt-pid}
rt_pid <- fdata %>% 
  group_by(pid, stim, compat) %>% 
  summarise(n = n(), 
            mean = mean(rt), 
            sd = sd(rt),
            sem = sd/sqrt(n))
rt_pid
```

group summary

```{r rt-group}
rt_group <- fdata %>% 
  group_by(stim, compat) %>% 
  summarise(n = n(), 
            mean = mean(rt), 
            sd = sd(rt),
            sem = sd/sqrt(length(unique(pid))))
rt_group
```

plots

density

```{r plot3.2}
p3.2 <- ggplot(fdata, 
               aes(x = rt, y=after_stat(count),
                   fill = compat)) +
  geom_density(alpha = 0.5) +
  scale_fill_brewer(palette = "Dark2") +
  facet_grid(~stim)
p3.2  
```

violin

```{r plot3.3}
p3.3 <- ggplot(fdata, 
               aes(x = compat, y = rt, fill = compat)) +
  geom_jitter(data = rt_pid, aes(y = mean), 
              position=position_jitterdodge(dodge.width =1), 
               alpha = 1, colour = "darkgrey") +
  geom_violin(alpha = 0.7, position=pd2) +
  geom_point(data = rt_group, aes(y = mean), size = 3) +
  geom_errorbar(data = rt_group,
                aes(y = mean, ymin = mean-sem, ymax = mean +sem),
                width=.2, position=pd2) +
  geom_line(data = rt_group, aes(y = mean, group=1)) +
  scale_fill_brewer(palette = "Dark2") +
  facet_grid(~stim) +
  ylim(250, 1250) +
  labs(title = "Reaction time")
p3.3  

ggsave("exp1/figures/rt_violin.png",
       width = 6, height = 8)
```

plot them together 

```{r plot-3.3}
p3.3 <- p3.1 | p3.2
p3.3
```

## section 4 ##

## calculate difference scores ##

this is on the summary per pid data

```{r diffs-pid}
diffs <- rt_pid %>%
  group_by(pid, stim) %>% 
  pivot_wider(id_cols = -c(n, sd, sem),
              names_from = "compat",
              values_from = "mean") %>%
  summarise(diff = incomp - compat) 
diffs
```

summary data

```{r diffs-group}
diffs_group <- diffs %>%
  group_by(stim) %>% 
  summarise(n = n(), 
            mean = mean(diff), 
            sd = sd(diff),
            sem = sd/sqrt(n))
diffs_group
```

density

```{r plot4.1}
p4.1 <- ggplot(diffs, 
               aes(x = diff, y=after_stat(count),
                   fill = stim)) +
  geom_density(alpha = 0.5) +
  scale_fill_brewer(palette = "Dark2") +
  labs(title = "Difference score") 
p4.1  

# ggsave("figures/diff_dens.png")
```

histogram

```{r plot4.2}
p4.2 <- ggplot(diffs, 
               aes(x = diff, fill = stim, colour = stim)) +
  geom_histogram(alpha = 0.5, position = "identity") +
  geom_rug() +
  scale_fill_brewer(palette = "Dark2") +
  scale_colour_brewer(palette = "Dark2") +
  labs(title = "Difference score") 
p4.2

# ggsave("figures/diff_hist.png")
```

violin

```{r plot4.3}
p4.3 <- ggplot(diffs, 
               aes(x = stim, y = diff, fill = stim)) +
  geom_jitter(position=position_jitterdodge(dodge.width =1), 
              alpha = 1, colour = "darkgrey") +
  geom_violin(alpha = 0.7, position=pd2) +
  geom_point(data = diffs_group, aes(y = mean), size=3) +
  geom_errorbar(data = diffs_group,
                aes(y = mean, ymin = mean-sem, ymax = mean+sem), 
                width=.2, position=pd2) +
  scale_fill_brewer(palette = "Dark2") +
  geom_line(data = diffs_group, aes(y = mean, group=1)) +
  labs(title = "Difference score") 
p4.3  

ggsave("exp1/figures/diff_violin.png",
       width = 6, height = 8)
```

ok, so this seems like a small effect in the expected direction for both types of
stimuli. 


let's try to display percentiles for the size of the difference score

using ggridges and with data points too

```{r plot4.4}
p4.4 <- ggplot(diffs, aes(x = diff, y = "", fill = after_stat(quantile))) +
  stat_density_ridges(
    geom = "density_ridges_gradient", calc_ecdf = TRUE,
    quantiles = 10, quantile_lines = TRUE, 
    jittered_points = TRUE, scale = 50,
    linewidth = 0.7, vline_color = "red",
    point_size = 1, point_alpha = 1, 
    position = position_raincloud(adjust_vlines = TRUE)
  ) +
  scale_fill_viridis_d(name = "Quantiles") +
  geom_vline(xintercept = 0, colour = "black", linetype = "dashed",
             linewidth = 1) +
  geom_vline(data = diffs_group,
             aes(xintercept = mean), colour = "turquoise", linetype = "dashed",
             linewidth = 1) +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 1)) +
  labs(title = "Difference score density by quantile",
       y = NULL,
       x = "difference score (ms)") +
  scale_x_continuous(breaks = seq(-150, 150, 25)) +
  facet_wrap(~stim, ncol = 1)
p4.4

ggsave("exp1/figures/diff_dens_q.png",
       width = 6, height = 8)
```

## section 5 ##

## save out some data ##

```{r save-data}
## all data
write_csv(fdata, "exp1/data/fdata.csv")
```

